---
title: "Basic Springboard Data Manipulation Exercise"
author: "Jonathan morken"
date: "April 7, 2016"
output: html_document
---

## Load the Data
```{r, warning=FALSE, message=FALSE}
setwd("C:/Users/jmorken/Documents/GitHub/Data Wrangling Projects/Basic Springboard Data Manipulation Exercise")

df <- read.csv("Data/refine_original.csv", sep = ",", header = TRUE)
df$company <- as.character(df$company)
df$company <- tolower(df$company)
df$Product.code...number <- as.character(df$Product.code...number)
df$address <- as.character(df$address)
df$city <- as.character(df$city)
df$country <- as.character(df$country)
df$name <- as.character(df$name)
```

## Clean up Brand Names
Clean up the 'company' column, so all of the misspellings of the brand names are standardized. For example, you can transform the values in the column to be: philips, akzo, van houten and unilever (all lowercase).
```{r, warning=FALSE, message=FALSE}
# Load the libraries
library(dplyr)
library(tidyr)

#Add Company Key
companies <- c("philips", "akzo", "van houten","unilever")
companies <- as.data.frame(companies)
companies$companies <- as.character(companies$companies)

#install.packages(stringdist)
library(stringdist)

# Map messy company character strings to standard versions.
d <- expand.grid(unique(df$company),companies$companies, KEEP.OUT.ATTRS = FALSE, stringsAsFactors = FALSE) # Create long form of original vs. key
names(d) <- c("company","b_name")  # Name the columns
d$dist <- stringdist(d$company,d$b_name, method="jw")  # Calculate fit 
d <-do.call(rbind, unname(by(d, d$company, function(x) x[x$dist == min(x$dist),])))  # Reduce to best fit

# Join created data frame with original, re-map company and get rid of b_name column.
df <- left_join(df, d, by = "company")
df$company <- df$b_name
df <- df[,c(1:6)]
```

## Separate product code and number
Separate the product code and product number into separate columns i.e. add two new columns called product_code and product_number, containing the product code and number respectively
```{r, warning=FALSE, message=FALSE}
df <- df %>%
  separate(Product.code...number, c("product_code", "product_number"), sep = "\\-")
```

## Add product categories
You learn that the product codes actually represent the following product categories:
*  p = Smartphone
*  v = TV
*  x = Laptop
*  q = Tablet

In order to make the data more readable, add a column with the product category for each record.
```{r, warning=FALSE, message=FALSE}
product_codes <- data.frame(product_code = c("p", "v", "x","q"), product_category = c("Smartphone","TV","Laptop","Tablet"))

df <- left_join(df,product_codes)
```

## Add full address for geocoding
You'd like to view the customer information on a map. In order to do that, the addresses need to be in a form that can be easily geocoded. Create a new column full_address that concatenates the three address fields (address, city, country), separated by commas.
```{r, warning=FALSE, message=FALSE}
df <- unite(df, full_address, address, city, country, sep = ",", remove = FALSE)
```

## Create dummy variables for company and product category
Both the company name and product category are categorical variables i.e. they take only a fixed set of values. In order to use them in further analysis you need to create dummy variables. Create dummy binary variables for each of them with the prefix company_ and product_ i.e.
*  Add four binary (1 or 0) columns for company
*  Add four binary (1 or 0) columns for product category

```{r, warning=FALSE, message=FALSE}
#company variables
df <- df %>%
  mutate(company = paste("company", company, sep = "_"), yesno = 1) %>%
  distinct %>%
  spread(company, yesno, fill = 0)
#category variables
df <- df %>%
  mutate(product_category = paste("product", product_category, sep = "_"), yesno = 1) %>%
  distinct %>%
  spread(product_category, yesno, fill = 0)
```

## Create final CSV
```{r, warning=FALSE, message=FALSE}
write.csv(df, file = "data/refine_clean.csv",row.names = FALSE)
```
